---
title: "Data processing from MIP to csv databases"
author: 
- name: "Jason_Rosado"
affiliation: Institut de Recherche pour le Développement
email: jason.rosado@ird.fr 
date: "`r Sys.Date()`"
output: 
  html_document:
    css: mystyle.css
    number_sections: yes
    toc: true
    toc_float: true
    code_folding: hide 
    code_download: false
    fig_width: 8
    fig_height: 6
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      tidy = TRUE,
                      fig.width = 8, 
                      fig.height = 6)

library(plyr)
library(table1)
library(tidyverse)
library(table1)
library(miplicorn)
library(ggplot2)
library(cowplot)
library(ggpubr)
library(here)
library(ComplexHeatmap)
library(heatmaply)
library(magick)
```

# Introduction

This R Markdown document was written by Jason Rosado using data generated during his visit to Jeff Bailey laboratory at Brown University, with help of Abebe Fola and Rebecca Crudale. This work is lead by Jerome Clain (IRD/ Universite Paris Cite) <https://umr-merit.ird.fr/en/axes-et-projets/meds/>, Sandrine Houze (IRD/CNR)<https://cnr-paludisme.fr/> and Jeff Bailey (Brown University) <https://baileylab.org/>. 

For more details on MIPs for Drug Resistance see <https://doi.org/10.1093/infdis/jiy223>.

DR2 MIP panel was used on 830 DNA samples (10 plates). The DR2 MIP panel covered 43 genes, from which 13 genes have been associated with drug resistance. 815 probes targeting different loci across resistance genes.

# Merging coverage and genotype MIP data
Merge tables of reference aminoacids, alternative aminoacids and coverage. 
To merge these data we use the built-in function in **miplicorn** <https://github.com/bailey-lab/miplicorn>

We are removing samples with less than 3 UMI coverage. Now we have: Number of individuals = 813; Mutation_name= 4109 ; Number of genes = 43. Genotype is classified as 'wild-type' or '0' if min_freq of alternative allele is <0; 'mutant'  or '2' if min_freq of alternative allele is '>=1'; 'mixed genotype' or 1 if  min_freq>'0' & min_freq<'1' 
```{r, echo=FALSE, message=FALSE}
#FILES PATH
ref_file <- (here("raw_data", "reference_AA_table.csv")) 
alt_file <- (here("raw_data","alternate_AA_table.csv"))
cov_file <- (here("raw_data","coverage_AA_table.csv"))

data_all_plates<- read_tbl_ref_alt_cov(ref_file,
                                alt_file,
                                cov_file) %>% # We filtered out the positive and negative contrlos included in the sequencing run
  filter(!sample %in% c("3D7-4000p-CNR-1","3D7-2000p-CNR-1", 
                        "3D7-1000p-CNR-1", "3D7-500p-CNR-1", "3D7-200p-CNR-1",
                        "3D7-100p-CNR-1", "3D7-50p-CNR-1", "3D7-10p-CNR-1",
                        "NTC-CNR-1", "NTP-CNR-1"), #coverage = minimun number of UMI reads
         coverage >=3 )  %>% 
  mutate(aa_change = convert_three(aa_change),  #this allows to convert to nomenclature of three amino acids
        min_freq = alt_umi_count/(alt_umi_count+ref_umi_count), #estimates the proportion of alternative allele in case of mixed genotype samples
        across(min_freq, ~replace(.,is.nan(.),0 )) ,
      #genotype is classified as 'wild-type' or '0' if min_freq of alternative allele is <0; 'mutant'  or '2' if min_freq of alternative allele is '>=1'; 'mixed genotype' or 1
        genotype = case_when(min_freq<="0" ~"0",
        min_freq>="1" ~ "2",
        min_freq>"0" & min_freq<"1" ~ "1")) 

```


```{r}
dir.create("intermediate_data")
dir.create("figures")
dir.create("tables")

write_rds(data_all_plates, "intermediate_data/MIP_data_all_plates.rds", "xz")
```

## Plot average coverage
First step is to check the coverage per barcode. The expected number of reads per sample is: 815 (probes) * 100 (ideal number of reads) = 81,500 reads; but we could also work with 815 *10 = 8,150 reads per sample. 
Some samples perform very poorly for one reason or another. This library have been repooled at a higher volume already.The samples with a total barcode count of 100 is considered low covered.
```{r, echo=FALSE}
plot_coverage<- plot_coverage(data_all_plates, group_by = gene) +
  ylab("mean coverage per sample")
print(plot_coverage)
ggsave("figures/plot_coverage.pdf", units="cm", width=20, 
       height=15, dpi=300)


```

##Mean coverage table
```{r}
table1(~  mean_coverage
     , data = plot_coverage$data)
```

```{r,echo=FALSE, message=FALSE}
merged_genotype <- data_all_plates %>%
  dplyr::select( mutation_name,sample,gene_id,gene, 
          aa_change,min_freq,genotype) %>%
  mutate(sample = as.numeric(sub("-CNR-1", '',sample))) %>%
  unique()

```

## Load genotype data: Include mutation frequency >1% and exclude duplicated mutation names
We started with 4109 'mutation names'. After excluding mutations with <1% frequency we ended up with 1201 mutations.

```{r, echo=FALSE, warning=FALSE}

#We modify the sample name in the genotypes table to match with the original database IC50_CNR
genotype_modified <- 
  data_all_plates %>%
  dplyr::select(mutation_name,sample,gene_id, gene, aa_change,min_freq,genotype) %>%
  mutate(sample = as.numeric(sub("-CNR-1", '',sample))) %>%
  #Most of duplicated mutation names have missing data
  #we are filtering the mutation names that are in the dataframe mutation frequency with >1% frequency
  rename(mutation = mutation_name) %>%
  group_by(sample, mutation) %>%
  distinct(sample, mutation, .keep_all = T) %>%
  dplyr::select(!c(gene,gene_id, aa_change, min_freq)) %>%
  mutate(genotype = as.integer(genotype)) %>%
  pivot_wider(names_from = mutation, values_from = genotype)

```


## Load CNR database
```{r, echo=FALSE, message=FALSE}

CNR_IC50_epidata <- read.csv(here("raw_data","CNR_IC50_epidata.csv")) 

```

## Load COI data
Loading estimated COI of samples  
```{r}
COI_data<- read.csv(here("raw_data","COI_calls.csv")) %>%
  rename(median_coi = median,
         sample =  sample_names) %>%
  dplyr::select(sample, median_coi)
```

## Merge MIP data, COI data and CNR database
At this point we have 1201'mutation name'. This is the reference database for prevalence and other analysis
```{r, echo=FALSE, message=FALSE}
#This is the reference database for prevalence and other analysis

IC50_genotype<- CNR_IC50_epidata %>%
  inner_join(genotype_modified, by = "sample") %>%
  inner_join(COI_data, by = "sample")
```


```{r, echo=F, eval=F}
write_rds(IC50_genotype,"intermediate_data/IC50_genotype.rds", "xz")

```

##Plot coverage heatmap
#Loading barcode counts
```{r}
barcode_file <- read.csv(here("raw_data","barcode_counts.csv")) %>%
  rename(sample_ID = MIP) %>%
  filter(!c(sample_ID =="Sample ID" | sample_ID =="Copy")) %>%
  pivot_longer(!sample_ID, names_to = "MIP", values_to = "UMI") %>%
  mutate(UMI = as.numeric(UMI),
         sample_ID = as.numeric(sub("-CNR-1", '',sample_ID))) %>%
  mutate(MIP = str_replace_all(MIP,"[\\.]|([\\.]?=[\\_])|([\\.]?=[\\_])","_")) %>%
  #removing mips that never went sequenced (they have Zero read counts)
filter(!c(MIP=="pfabcI3_S0_Sub0_mip35" |MIP=="pfv_atpase_S0_Sub0_mip4" | MIP=="pfpi4k_S0_Sub0_mip19" |  MIP=="pfpi4k_S0_Sub0_mip32"))
   
```

# Filtering only samples with good sequencing data
```{r}
barcode_file <- barcode_file %>%
  filter(sample_ID %in% IC50_genotype$sample)
```

# Loading MIPs used for this sequencing run
```{r}
MIPs <- read.csv(here("raw_data","all_haplotypes.csv"))%>%
  dplyr::select(MIP, Gene) %>%
  mutate(MIP = str_replace_all(MIP,"[\\-]|([\\-]?=[\\_])|([\\-]?=[\\_])","_"))%>%
  unique() %>%
  rename(gene = Gene) 
```

#Creating parasitemia classes from database
```{r, warning=FALSE}
sample_parasitemia<- IC50_genotype %>%
  select(sample, parasitemia) %>%
  rename(sample_ID = sample) %>%
  mutate(parasitemia =  as.numeric(parasitemia),
         parasitemia = cut_number(parasitemia, 6))
  
```

#Coverage Heatmap/Probes/Samples
```{r}
MIP_key_genes<-MIPs %>%
  mutate(gene = case_when(gene=="dhfr-ts" ~ "dhfr-ts",
                          gene=="mdr1" ~ "mdr1",
                          gene =="crt" ~ "crt",
                          gene =="dhps" ~ "dhps",
                          gene =="k13" ~ "k13",
                          TRUE ~ "others"
                          ))
```


```{r}
pheat_data <- barcode_file %>%
                     group_by(sample_ID, MIP) %>%
                     summarise(
                       count = log10(sum(UMI) + 1)
                     ) %>%
  filter(MIP %in% MIP_key_genes$MIP) %>%
                     ungroup() %>%
                     pivot_wider(
                       names_from = MIP,
                       values_from = count
                     ) %>% as.data.frame()
  
  rownames(pheat_data) <- pheat_data$MIP
  
  # Annotation color
  annotation2 <- sample_parasitemia %>% select(sample_ID, parasitemia)
  rownames(annotation2) <- annotation2$sample_ID
  rowname_anno2 <- rownames(annotation2)
  annotation2 <- data.frame(parasitemia = annotation2$parasitemia)
  rownames(annotation2) <- rowname_anno2

  
  
  pheat_data <- pheat_data[, -1]
  annotation <- data.frame(MIP = colnames(pheat_data))
  annotation <- annotation %>%
    left_join(
      (MIP_key_genes %>%
         select(MIP, gene)),
      by = join_by(MIP)
    ) %>%
    filter(!is.na(gene)) %>%
    as.data.frame()
  rownames(annotation) <- annotation$MIP
  rowname_anno <- rownames(annotation)
  annotation <- data.frame(gene = annotation$gene)
  rownames(annotation) <- rowname_anno
 
  
  # Color 
  suppressMessages(pal_info <- annotation %>%
                     select(gene) %>%
                     distinct() %>%
                     mutate(colour = scales::hue_pal()(n())) %>%
                     with(setNames(colour, gene)))
  
  pal_info <- list(
    gene = pal_info
  )
  
  suppressMessages(parasitemia <- annotation2 %>%
                     select(parasitemia) %>%
                     distinct() %>%
                     mutate(colour = scales::brewer_pal()(n())) %>%
                     with(setNames(colour, annotation2)))
  
  parasitemia <- list(
    parasitemia = parasitemia
  )
  
  pheat_data[is.na(pheat_data)] <- 0
  col <- colorRampPalette(rev(RColorBrewer::brewer.pal(n = 7, name = "RdYlBu")))(100)
  
  col <- c("#BEBEBE", col)
 
  
  pdf("figures/MIP_UMI.pdf", width=8, height=7)

  
  ht1 = ComplexHeatmap::pheatmap(t(pheat_data),
                           name = "Log10(UMI + 1)",
                           row_title = "MIP",
                           color = col,
                           column_title = "sample",
                           show_rownames = FALSE,
                           show_colnames = FALSE,
                           row_split = annotation$gene,
                           annotation_colors = pal_info,
                           annotation_row = annotation,
                           annotation_col = annotation2,
                           cluster_rows = FALSE,
                           cluster_cols = TRUE,
                           use_raster = TRUE,
                           raster_by_magick = TRUE,
                           raster_magick_filter = "Bartlett"
  )
  
  draw(ht1, row_sub_title_side = "left")
  dev.off()


```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
MOI<- IC50_genotype %>%
  dplyr::select(median_coi) %>%
  ggplot(aes(median_coi)) +
  geom_bar(position = position_dodge2(width = 1, preserve = "single")) + 
  ggtitle("") +
  xlab("COI") + ylab("Frequency") +
  scale_fill_brewer(palette="Paired") + scale_x_continuous(breaks = seq(1, 12, 1))+
  theme_classic(base_size = 12)  

print(MOI)
ggsave("figures/COI_dist.pdf", units="cm", width=10, height=10, dpi=300)

parasite_hist<- IC50_genotype %>%
  mutate(parasitemia = as.numeric(parasitemia)) %>%
  dplyr::select(parasitemia) %>%
  ggplot(aes(parasitemia)) +
  geom_histogram(
    position = position_dodge2(width = 0.9, preserve = "single")
    ) +
  ggtitle("") +
  xlab("Parasitemia (% iRBC)") + ylab("Frequency") +
  scale_fill_brewer(palette="Paired") +
  theme_classic(base_size = 12)  


print(parasite_hist)
ggsave("figures/parasite_hist.pdf", units="cm", width=10, height=10, dpi=300)


```

```{r echo=F, eval=F}
save.image("./intermediate_data/data_cleaning.RData")
```





