---
title: "Half-maximal inhibitory concentration IC50"
author: 
- name: "Jason_Rosado"
affiliation: Institut de Recherche pour le Développement
email: jason.rosado@ird.fr 
date: "`r Sys.Date()`"
output: 
  html_document:
    css: mystyle.css
    number_sections: yes
    toc: true
    toc_float: true
    code_folding: hide 
    code_download: false
    fig_width: 8
    fig_height: 6
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      tidy = TRUE,
                      fig.width = 8, 
                      fig.height = 6)

library(plyr)
library(table1)
library(tidyverse)
library(table1)
library(ggplot2)
library(cowplot)
library(grid)
library(jtools)
library(ggpubr)
library(corrplot)
library(ggcorrplot)
library(forcats)
library(here)
library(tableone)
```

#Loading data from previous runs
```{r, include=FALSE}
IC50_genotype <- readRDS(here("intermediate_data", "IC50_genotype.rds"))

WAF_CAF_class <- readRDS(here("intermediate_data", "WAF_CAF_class.rds")) 
```


# Half-maximal inhibitory concentration (IC50) for six antimalarial drugs

## Drug correlation matrix
```{r, echo=FALSE, message=FALSE}

data_drug<-IC50_genotype %>%
  dplyr::select(DHA, Lumefantrine, Mefloquine, MDAQ, chloroquine,piperaquine) %>%
  #data is log10 transform to get normal distribution
  mutate(DHA = log10(DHA), Lumefantrine =log10(Lumefantrine), Mefloquine = log10(Mefloquine), MDAQ = log10(MDAQ), chloroquine = log10(chloroquine),piperaquine = log10(piperaquine)) %>%
  rename(Dihydroartemisinin = DHA,
         Chloroquine = chloroquine,
         Piperaquine = piperaquine,
         Monodesethylamodiaquine = MDAQ)

#we use pearson correlation as the IC50 has now a normal distribution
ct2=cor(data_drug, method = "pearson")


p.mat <- cor_pmat(data_drug, "pearson", alternative = "two.sided")

```


```{r, echo=FALSE, message=FALSE}
drug_correlation<- ggcorrplot(ct2, p.mat= p.mat,hc.order = TRUE, type = "lower", outline.col = "black",  
           lab = TRUE, insig = "blank", ggtheme = ggplot2::theme_classic(base_size = 12)) 
print(drug_correlation)
ggsave("figures/drug_correlation.pdf", units="cm", width=19, height=15, dpi=300)

```
# Drug susceptibility by year
```{r}
IC50_year<- IC50_genotype %>%
  dplyr::select(year, DHA, Lumefantrine, Mefloquine, MDAQ, chloroquine,piperaquine) %>%
  mutate(year = as.character(year),
  year = case_when(year=="2020" ~ "2020-21", 
                          year=="2021"  ~ "2020-21", 
                        year=="2022" ~ "2022-23", 
                          year=="2023" ~ "2022-23", 
                          TRUE~ year))

IC50_year_table<- table1(~  DHA+ 
         Lumefantrine + Mefloquine + MDAQ + chloroquine +
        piperaquine | year
     , data = IC50_year,render.continuous = c("Median [Q1, Q3]") )

write.csv(IC50_year_table, "tables/IC50_year_table.csv", col.names = F)

```


```{r, echo=FALSE, message=FALSE}
DHA<- IC50_genotype %>%
  dplyr::select(year, DHA) %>%
  rename(Dihydroartemisinin = DHA) %>%
  mutate(year = as.character(year),
  year = case_when(year=="2020" ~ "2020-21", 
                          year=="2021"  ~ "2020-21", 
                        year=="2022" ~ "2022-23", 
                          year=="2023" ~ "2022-23", 
                          TRUE~ year))%>%
  group_by(year) %>%
  ggplot(aes(year, Dihydroartemisinin,fill =year)) + scale_y_log10() +
   geom_boxplot() + scale_fill_brewer(palette = "Pastel1") +
  geom_hline(yintercept = 6, linetype="dashed",
               color = "red", size=1) +
  ylab("IC50(nM)") + xlab("")+
ggtitle("Dihydroartemisinin") + theme_classic() +theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_blank(), axis.ticks.x = element_blank() )

Lumefantrine<- IC50_genotype %>%
  dplyr::select(year, Lumefantrine) %>%
  mutate(year = as.character(year),
  year = case_when(year=="2020" ~ "2020-21", 
                          year=="2021"  ~ "2020-21", 
                        year=="2022" ~ "2022-23", 
                          year=="2023" ~ "2022-23", 
                          TRUE~ year))%>%
  group_by(year) %>%
  ggplot(aes(year, Lumefantrine, fill =year )) + scale_y_log10() +
  geom_boxplot() + scale_fill_brewer(palette = "Pastel1") +
  geom_hline(yintercept = 150, linetype="dashed",
                color = "red", size=1) +
  ylab("") + xlab("")+
ggtitle("Lumefantrine") + theme_classic() +theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_blank(), axis.ticks.x = element_blank() )

Mefloquine<- IC50_genotype %>%
  dplyr::select(year, Mefloquine) %>%
  mutate(year = as.character(year),
  year = case_when(year=="2020" ~ "2020-21", 
                          year=="2021"  ~ "2020-21", 
                        year=="2022" ~ "2022-23", 
                          year=="2023" ~ "2022-23", 
                          TRUE~ year))%>%
  group_by(year) %>%
  ggplot(aes(year, Mefloquine, fill =year )) + scale_y_log10() +
  geom_boxplot() + scale_fill_brewer(palette = "Pastel1") +
  geom_hline(yintercept = 30, linetype="dashed",
                color = "red", size=1) +
  ylab("") + xlab("")+
ggtitle("Mefloquine") + theme_classic() +theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_blank(), axis.ticks.x = element_blank() )

Monodesethylamodiaquine<- IC50_genotype %>%
  dplyr::select(year, MDAQ) %>%
  rename(Monodesethylamodiaquine = MDAQ) %>%
  mutate(year = as.character(year),
  year = case_when(year=="2020" ~ "2020-21", 
                          year=="2021"  ~ "2020-21", 
                        year=="2022" ~ "2022-23", 
                          year=="2023" ~ "2022-23", 
                          TRUE~ year))%>%
  group_by(year) %>%
  ggplot(aes(year, Monodesethylamodiaquine, fill =year )) + scale_y_log10() +
  geom_boxplot() + scale_fill_brewer(palette = "Pastel1") +
  geom_hline(yintercept = 80, linetype="dashed",
                color = "red", size=1) +
  ylab("IC50(nM)") + xlab("")+
ggtitle("Monodesethylamodiaquine") + theme_classic() +theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_blank(), axis.ticks.x = element_blank() )

Chloroquine<- IC50_genotype %>%
  dplyr::select(year, chloroquine) %>%
  rename(Chloroquine = chloroquine) %>%
  mutate(year = as.character(year),
  year = case_when(year=="2020" ~ "2020-21", 
                          year=="2021"  ~ "2020-21", 
                        year=="2022" ~ "2022-23", 
                          year=="2023" ~ "2022-23", 
                          TRUE~ year))%>%
  group_by(year) %>%
  ggplot(aes(year, Chloroquine, fill =year )) + scale_y_log10() +
  geom_boxplot() + scale_fill_brewer(palette = "Pastel1") +
  geom_hline(yintercept = 100, linetype="dashed",
                color = "red", size=1) +
  ylab("") + xlab("")+
ggtitle("Chloroquine") + theme_classic() +theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_blank(), axis.ticks.x = element_blank() )

Piperaquine<- IC50_genotype %>%
  dplyr::select(year, piperaquine) %>%
  rename(Piperaquine = piperaquine) %>%
  mutate(year = as.character(year),
  year = case_when(year=="2020" ~ "2020-21", 
                          year=="2021"  ~ "2020-21", 
                        year=="2022" ~ "2022-23", 
                          year=="2023" ~ "2022-23", 
                          TRUE~ year))%>%
  group_by(year) %>%
  ggplot(aes(year, Piperaquine, fill =year )) + scale_y_log10() +
  geom_boxplot() + scale_fill_brewer(palette = "Pastel1") +
  geom_hline(yintercept = 135, linetype="dashed",
                color = "red", size=1) +
  ylab("") + xlab("")+
ggtitle("Piperaquine") + theme_classic() +theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_blank(), axis.ticks.x = element_blank() )

drugs_dist<- ggarrange(DHA, Lumefantrine, Mefloquine,Monodesethylamodiaquine,
          Chloroquine,Piperaquine,
                              ncol=3, nrow=2, common.legend = TRUE, legend="bottom")
print(drugs_dist)

ggsave("figures/drug_dist_yearly.pdf", units="cm", width=22, height=14, dpi=300)

  
```


## Calculate Kendall's tau and its associated p-value
```{r}
library(Kendall)
#2020 represents "2020-21" period  
#2021 represents "2022-23" period

DHA_trend<-c(1.10,1.16,1.05,0.940,1.28,	1.42) #These are IC50 median per year
year<- c(2016, 2017, 2018, 2019, 2020, 2021) 
tau_result_DHA <- cor.test(DHA_trend, year, method = "kendall")

LMF_trend<-c(15.5,14.7,11.3,18.0,24.8 ,25.3)
tau_result_LMF <- cor.test(LMF_trend, year, method = "kendall")

MFQ_trend<-c(33.3 ,	40.5 ,	23.7 ,	20.1,	40.7 ,	30.6) 
tau_result_MFQ <- cor.test(MFQ_trend, year, method = "kendall")

MDAQ_trend<-c(19.7,	19.0 , 25.3, 35.0, 36.9, 34.8) 
tau_result_MDAQ <- cor.test(MDAQ_trend, year, method = "kendall")

CQ_trend<-c(24.3,27.3, 29.8,28.7, 29.8, 19.3) 
tau_result_CQ <- cor.test(CQ_trend, year, method = "kendall")

PPQ_trend<-c(18.0,19.6, 16.1, 19.1, 25.4, 19.5) 
tau_result_PPQ <- cor.test(PPQ_trend, year, method = "kendall")
```


```{r}
# Install if needed
install.packages("trend")  # For Mann-Kendall test
install.packages("dplyr")   # For data wrangling
install.packages("ggplot2") # For plotting trends
```

## Calculate Kendall's tau (aggregate data) and its associated p-value with Trend package
```{r}

library(trend)

DHA_trend<-c(1.10,1.16,1.05,0.940,1.28,	1.42) #These are IC50 median per year
year<- c(2016, 2017, 2018, 2019, 2020, 2021) 
mk_result_DHA <- mk.test(DHA_trend)

LMF_trend<-c(15.5,14.7,11.3,18.0,24.8 ,25.3)
mk_result_LMF <- mk.test(LMF_trend)

MFQ_trend<-c(33.3 ,	40.5 ,	23.7 ,	20.1,	40.7 ,	30.6) 
mk_result_MFQ <- mk.test(MFQ_trend)

MDAQ_trend<-c(19.7,	19.0 , 25.3, 35.0, 36.9, 34.8) 
mk_result_MDAQ <- mk.test(MDAQ_trend)

CQ_trend<-c(24.3,27.3, 29.8,28.7, 29.8, 19.3) 
mk_result_CQ <- mk.test(CQ_trend)

PPQ_trend<-c(18.0,19.6, 16.1, 19.1, 25.4, 19.5) 
mk_result_PPQ <- mk.test(PPQ_trend)
```

# IC50 data stratified into 2016-17 vs 2018-23 periods
```{r}
IC50_before2017_after2017 <- IC50_genotype %>%
  dplyr::select(year, DHA, Lumefantrine, Mefloquine, MDAQ, chloroquine,piperaquine) %>%
  mutate( year = ifelse(year<=2017, "2016-17","2018-23"))

```

## Distribution of resistant isolates in 2016-17 versus 2018-23
-Pascual- Aurélie- Marilyn Madamet- Sébastien Briolant- Tiphaine Gaillard- Rémy Amalvict- Nicolas Benoit- Dominique Travers- Bruno Pradines- and French National Reference Centre for Imported Malaria Study Group. 2015. “Multinormal in Vitro Distribution of Plasmodium Falciparum Susceptibility to Piperaquine and Pyronaridine.” Malaria Journal 14 (February): 49.
-Kaddouri- Halima- Abdoulaye Djimdé- Souleymane Dama- Aly Kodio- Mamadou Tekete- Véronique Hubert- Aminatou Koné- et al. 2008. “Baseline in Vitro Efficacy of ACT Component Drugs on Plasmodium Falciparum Clinical Isolates from Mali.” International Journal for Parasitology 38 (7): 791–98.
```{r}
Category_drugs<- IC50_genotype %>%
  dplyr::select(year, country ,DHA, Lumefantrine, Mefloquine, MDAQ, chloroquine,piperaquine) %>%
  rename(Dihydroartemisinin = DHA,
         Chloroquine = chloroquine,
         Piperaquine = piperaquine,
         Monodesethylamodiaquine = MDAQ) %>%
  mutate(year = as.factor(year),
  LMF_cat = ifelse(Lumefantrine>150, "R","S"), #Kaddouri 2008
  DHA_cat = ifelse(Dihydroartemisinin>10,"R","S"), #Kaddouri 2008
  MFQ_cat = ifelse(Mefloquine>30, "R", "S"), #Pradines 2002
  PPQ_cat = ifelse(Piperaquine>135, "R", "S"), #Pradines 2002   
  CQ_cat = ifelse(Chloroquine>100,"R", "S"),#Kaddouri 2008
  MDAQ_cat = ifelse(Monodesethylamodiaquine>80,"R", "S")) #Kaddouri 2008
 

LUM_MFQ_scatter<-
ggplot(Category_drugs, aes(x=Lumefantrine, 
                                      y = Mefloquine, colour = year)) + 
         geom_point() + 
   geom_hline(yintercept = 30, linetype="dashed", 
                 color = "red", linewidth=1) + 
  geom_vline(xintercept = 150, linetype="dashed", 
                 color = "blue", linewidth=1) +
  theme_light(base_size = 14)
print(LUM_MFQ_scatter)
ggsave("figures/LUM_MFQ_scatter.pdf", units="cm", width=22, height=14, dpi=300)


LUM_DHA_scatter<-
ggplot(Category_drugs, aes(x= Lumefantrine, 
                                      y = Dihydroartemisinin, colour = year)) + 
         geom_point() + 
   geom_hline(yintercept = 10, linetype="dashed", 
                 color = "red", linewidth=1) + 
  geom_vline(xintercept = 150, linetype="dashed", 
                 color = "blue", linewidth=1)+
  theme_light(base_size = 16) + theme(axis.title= element_text(size = 18))
              
print(LUM_DHA_scatter)
ggsave("figures/LUM_DHA_scatter.pdf", units="cm", width=22, height=14, dpi=300)

MFQ_DHA_scatter<-
ggplot(Category_drugs, aes(x= Mefloquine, 
                                      y = Dihydroartemisinin, colour = year)) + 
         geom_point() + 
   geom_hline(yintercept = 10, linetype="dashed", 
                 color = "red", linewidth=1) + 
  geom_vline(xintercept = 30, linetype="dashed", 
                 color = "blue", linewidth=1)+
  theme_light(base_size = 16) + theme(axis.title= element_text(size = 18))
print(MFQ_DHA_scatter)
ggsave("figures/MFQ_DHA_scatter.pdf", units="cm", width=22, height=14, dpi=300)


PPQ_DHA_scatter<-
ggplot(Category_drugs, aes(x= Piperaquine, 
                                      y = Dihydroartemisinin, colour = year)) + 
         geom_point() + 
   geom_hline(yintercept = 10, linetype="dashed", 
                 color = "red", linewidth=1) + 
  geom_vline(xintercept = 135, linetype="dashed", 
                 color = "blue", linewidth=1)+
  theme_light(base_size = 16) + theme(axis.title= element_text(size = 18))
print(PPQ_DHA_scatter)
ggsave("figures/PPQ_DHA_scatter.pdf", units="cm", width=22, height=14, dpi=300)

MDAQ_DHA_scatter<-
ggplot(Category_drugs, aes(x= Monodesethylamodiaquine, 
                                      y = Dihydroartemisinin, colour = year)) + 
         geom_point() + 
   geom_hline(yintercept = 10, linetype="dashed", 
                 color = "red", linewidth=1) + 
  geom_vline(xintercept = 80, linetype="dashed", 
                 color = "blue", linewidth=1)  +
  theme_light(base_size = 16) + theme(axis.title= element_text(size = 18))

print(MDAQ_DHA_scatter)
ggsave("figures/MDAQ_DHA_scatter.pdf", units="cm", width=22, height=14, dpi=300)


CQ_DHA_scatter<-
ggplot(Category_drugs, aes(x= Chloroquine, 
                                      y = Dihydroartemisinin, colour = year)) + 
         geom_point() + 
   geom_hline(yintercept = 10, linetype="dashed", 
                 color = "red", linewidth=1) + 
  geom_vline(xintercept = 100, linetype="dashed", 
                 color = "blue", linewidth=1) +
  theme_light(base_size = 16) + theme(axis.title= element_text(size = 18))
print(CQ_DHA_scatter)
ggsave("figures/CQ_DHA_scatter.pdf", units="cm", width=22, height=14, dpi=300)

```

# IC50 distribution in 2016-17 vs 2018-23 period
```{r}
DHA_2016_2023_group<- IC50_before2017_after2017 %>%
  dplyr::select(year, DHA) %>%
  rename(Dihydroartemisinin = DHA) %>%
  mutate(year = as.character(year))%>%
  group_by(year) %>%
  ggplot(aes(year, Dihydroartemisinin,fill =year)) + scale_y_log10() +
   geom_boxplot() + scale_fill_brewer(palette = "Pastel1") +
  geom_hline(yintercept = 10, linetype="dashed", 
                color = "red", linewidth=1) + 
  ylab("IC50(nM)") + xlab("")+
ggtitle("Dihydroartemisinin") + theme_classic() +theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_blank(), axis.ticks.x = element_blank() ) +
  stat_compare_means(aes(group = year), label = "p.format", size = 4) #+
  #stat_summary(fun.data=f, geom = "text", vjust=-0.5, col ="black", size = 2.5)

Lumefantrine_2016_2023_group<- IC50_before2017_after2017 %>%
  dplyr::select(year, Lumefantrine) %>%
  mutate(year = as.character(year))%>%
  group_by(year) %>%
  ggplot(aes(year, Lumefantrine, fill =year )) + scale_y_log10() +
  geom_boxplot() + scale_fill_brewer(palette = "Pastel1") +
  geom_hline(yintercept = 150, linetype="dashed",
                color = "red", linewidth=1) +
  ylab("") + xlab("")+
ggtitle("Lumefantrine") + theme_classic() +theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_blank(), axis.ticks.x = element_blank() ) +
  stat_compare_means(aes(group = year), label = "p.format", size = 4)

Mefloquine_2016_2023_group<- IC50_before2017_after2017 %>%
  dplyr::select(year, Mefloquine) %>%
  mutate(year = as.character(year))%>%
  group_by(year) %>%
  ggplot(aes(year, Mefloquine, fill =year )) + scale_y_log10() +
  geom_boxplot() + scale_fill_brewer(palette = "Pastel1") +
  geom_hline(yintercept = 30, linetype="dashed",
                color = "red", linewidth=1) +
  ylab("") + xlab("")+
ggtitle("Mefloquine") + theme_classic() +theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_blank(), axis.ticks.x = element_blank() )+
  stat_compare_means(aes(group = year), label = "p.format", size = 4)

Monodesethylamodiaquine_2016_2023_group<- IC50_before2017_after2017 %>%
  dplyr::select(year, MDAQ) %>%
  rename(Monodesethylamodiaquine = MDAQ) %>%
  mutate(year = as.character(year))%>%
  group_by(year) %>%
  ggplot(aes(year, Monodesethylamodiaquine, fill =year )) + scale_y_log10() +
  geom_boxplot() + scale_fill_brewer(palette = "Pastel1") +
  geom_hline(yintercept = 80, linetype="dashed",
                color = "red", linewidth=1) +
  ylab("IC50(nM)") + xlab("")+
ggtitle("Monodesethylamodiaquine") + theme_classic() +theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_blank(), axis.ticks.x = element_blank() )+
  stat_compare_means(aes(group = year), label = "p.format", size = 4)

Chloroquine_2016_2023_group<- IC50_before2017_after2017 %>%
  dplyr::select(year, chloroquine) %>%
  rename(Chloroquine = chloroquine) %>%
  mutate(year = as.character(year))%>%
  group_by(year) %>%
  ggplot(aes(year, Chloroquine, fill =year )) + scale_y_log10() +
  geom_boxplot() + scale_fill_brewer(palette = "Pastel1") +
  geom_hline(yintercept = 100, linetype="dashed",
                color = "red", linewidth=1) +
  ylab("") + xlab("")+
ggtitle("Chloroquine") + theme_classic() +theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_blank(), axis.ticks.x = element_blank() ) +
  stat_compare_means(aes(group = year), label = "p.format", size = 4)

Piperaquine_2016_2023_group<- IC50_before2017_after2017 %>%
  dplyr::select(year, piperaquine) %>%
  rename(Piperaquine = piperaquine) %>%
  mutate(year = as.character(year))%>%
  group_by(year) %>%
  ggplot(aes(year, Piperaquine, fill =year )) + scale_y_log10() +
  geom_boxplot() + scale_fill_brewer(palette = "Pastel1") +
  geom_hline(yintercept = 135, linetype="dashed",
                color = "red", linewidth=1) +
  ylab("") + xlab("")+
ggtitle("Piperaquine") + theme_classic() +theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_blank(), axis.ticks.x = element_blank() ) +
  stat_compare_means(aes(group = year), label = "p.format", size = 4)

drugs_dist__2016_2023_group<- ggarrange(DHA_2016_2023_group, Lumefantrine_2016_2023_group, Mefloquine_2016_2023_group,Monodesethylamodiaquine_2016_2023_group,
          Chloroquine_2016_2023_group,Piperaquine_2016_2023_group,
                              ncol=3, nrow=2, common.legend = TRUE, legend="bottom")
ggsave("figures/drug_dist_2016_2023_group_cut_off.pdf", units="cm", width=22, height=14, dpi=300)
```

## Stats for IC50 distribution
```{r}
IC50_table<- table1(~  DHA+ 
         Lumefantrine + Mefloquine + MDAQ + chloroquine +
        piperaquine
     , data = IC50_genotype,render.continuous = c("Median [Q1, Q3]") )

write.csv(IC50_table, "tables/IC50_table.csv", col.names = F)


```

# Drug susceptibility by African region
```{r}

############################
#                          #
# WAF AND CAF REGIONS IC50 #
#                          #
############################

WAF_CAF_countries_IC50<- IC50_genotype %>%
  dplyr::select(sample, country, year,Lumefantrine,DHA,Mefloquine,MDAQ,
                chloroquine,piperaquine) %>%
   inner_join(WAF_CAF_class %>% dplyr::select(country,region), by = c("country")) 
 #%>%  mutate( year = ifelse(year<=2017, "2016-17","2018-23"))
  
WAF_CAF_IC_countries_table<- table1(~  Lumefantrine + DHA + Mefloquine + MDAQ +
                chloroquine + piperaquine  |region 
     , data = WAF_CAF_countries_IC50,render.continuous = c("Median [Q1, Q3]") )

write.csv(WAF_CAF_IC_countries_table, "tables/WAF_CAF_countries_IC50_table.csv", row.names = F)

```

## Comparison of IC50 between Central and West African countries and adjusted pvalues
```{r}
WAF_CAF_IC_region_year_stats<- CreateTableOne(vars = c("Lumefantrine", "DHA" , "Mefloquine" , "MDAQ" ,"chloroquine" , "piperaquine"), strata = c("region"), data = WAF_CAF_countries_IC50)

print(WAF_CAF_IC_region_year_stats,showAllLevels = TRUE,
      nonnormal = TRUE  )

#Wilcoxon test
wilcox.test(WAF_CAF_countries_IC50$Lumefantrine
            [WAF_CAF_countries_IC50$region=="CAF"], WAF_CAF_countries_IC50$Lumefantrine[WAF_CAF_countries_IC50$region=="WAF"], data= WAF_CAF_countries_IC50)

wilcox.test(WAF_CAF_countries_IC50$MDAQ 
            [WAF_CAF_countries_IC50$region=="CAF"], WAF_CAF_countries_IC50$MDAQ [WAF_CAF_countries_IC50$region=="WAF"], data= WAF_CAF_countries_IC50)

wilcox.test(WAF_CAF_countries_IC50$piperaquine 
            [WAF_CAF_countries_IC50$region=="CAF"], WAF_CAF_countries_IC50$piperaquine [WAF_CAF_countries_IC50$region=="WAF"], data= WAF_CAF_countries_IC50)

wilcox.test(WAF_CAF_countries_IC50$chloroquine  
            [WAF_CAF_countries_IC50$region=="CAF"], WAF_CAF_countries_IC50$chloroquine  [WAF_CAF_countries_IC50$region=="WAF"], data= WAF_CAF_countries_IC50)

wilcox.test(WAF_CAF_countries_IC50$DHA  
            [WAF_CAF_countries_IC50$region=="CAF"], WAF_CAF_countries_IC50$DHA   [WAF_CAF_countries_IC50$region=="WAF"], data= WAF_CAF_countries_IC50)

wilcox.test(WAF_CAF_countries_IC50$Mefloquine  
            [WAF_CAF_countries_IC50$region=="CAF"], WAF_CAF_countries_IC50$Mefloquine   [WAF_CAF_countries_IC50$region=="WAF"], data= WAF_CAF_countries_IC50)

#pvalues extracted by Wilcoxon test 
pvalues_regions<- c(0.7694 , 0.2418 , 0.0002728, 0.3782 , 0.1191, 0.7383)

p.adjust(pvalues_regions,method="BH")

#0.7694000 0.4836000 0.0016368 0.5673000 0.3573000 0.7694000 #adjusted P-values
```
## Stats for resistant isolates defined by cut-offs 
```{r}
Category_drugs_table<- table1(~  LMF_cat+ 
         DHA_cat + MFQ_cat + PPQ_cat + CQ_cat +
        MDAQ_cat 
     , data = Category_drugs)

write.csv(Category_drugs_table, "tables/Category_drugs_table.csv", col.names = F)

```

```{r echo=F, eval=F}
save.image("./intermediate_data/IC50_analysis_revised.RData")
```



