---
title: "Prevalence of validated antimalarial resistance mutations"
author: 
- name: "Jason_Rosado"
affiliation: Institut de Recherche pour le Développement
email: jason.rosado@ird.fr 
date: "`r Sys.Date()`"
output: 
  html_document:
    css: mystyle.css
    number_sections: yes
    toc: true
    toc_float: true
    code_folding: hide 
    code_download: false
    fig_width: 8
    fig_height: 6
    theme: cosmo
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      tidy = TRUE,
                      fig.width = 8, 
                      fig.height = 6)

library(plyr)
library(table1)
library(tidyverse)
library(miplicorn)
library(ggplot2)
library(cowplot)
library(grid)
library(jtools)
library(skimr)
library(ggpubr)
library(forcats)
library(here)
library(tableone)
library(writexl)
```

#Loading data from previous runs
```{r, include=FALSE}
IC50_genotype <- readRDS(here("intermediate_data", "IC50_genotype.rds"))
data_all_plates <- readRDS(here("intermediate_data", "MIP_data_all_plates.rds")) 
```

## Editing genotype codes to run SNPassoc.
We are including ALL monoclonal and polyclonal samples for this exploratory analysis. Excluding duplicated mutation names. For SNPassoc Format: Homozygous should be coded as 0 (wt), heterozygous as 1 and mutant homozygous as 2, and missing are coded as NA 
```{r, echo=FALSE, message=FALSE}
#We don't remove samples or mutation names, n = 830/Mutation name= 4110

long_format_genotype_data<- IC50_genotype %>%
  select(!c(palu_id, sample_date, species, parasitemia, year, sample_id2, 
            country, gender, age, DHA, Lumefantrine, Mefloquine,MDAQ,
            chloroquine,piperaquine, median_coi)) %>%
  pivot_longer(!c(sample), names_to = "mutation_name", values_to = "genotype") %>% 
  #removing duplicated sample id
  group_by(sample, mutation_name) %>%
  distinct(sample, mutation_name, .keep_all = T) 

```

## Check for duplicates: No duplicates
```{r, echo=FALSE, message=FALSE}
   long_format_genotype_data%>%
   dplyr::group_by(sample,mutation_name) %>%
   dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
   dplyr::filter(n > 1L) 

```


### We are removing mutation names with percentages of missing value >= 60%.
```{r, echo=FALSE,message=FALSE}
genotype_format_filtered<-
  long_format_genotype_data %>% 
  #We don't want duplicated sample and mutation name
  unique() %>%
  #Change format to be able to count for NA per mutation
  pivot_wider(names_from = sample, values_from = genotype) %>% 
  rowwise() %>%
  #We count the percentage of NA per mutation names 
  mutate(NA_percent = sum(is.na(across(everything())))/length(.),
         NAs_s = sum(is.na(across(everything())))) %>%
  #We remove mutation names with missing percentage >= 60%
  filter(NA_percent <=0.30) %>%
  #We are removing NAs_S and NA_percent as we don't need them anymore
  dplyr::select(!c(NA_percent,NAs_s)) %>%
  pivot_longer(cols = !mutation_name) %>%
  pivot_wider(names_from = "mutation_name", values_from = "value") %>%
    rename(sample = name)
  
#We have now 805 (before it was 813) unique samples and 3491 mutation names  
```


### Selecting IC50 phenotypic data
Only those sample names from upstream selection (n = 813). 
```{r, echo=FALSE,message=FALSE}
phenotypic_data_cont<- IC50_genotype%>%
  select(sample, DHA, Lumefantrine, Mefloquine,MDAQ,chloroquine,piperaquine,year,country,median_coi)

```

### Formatting genotype and phenotype data for prevalence analysis
Only those sample names from upstream selection (n = 805)
```{r, echo=FALSE,message=FALSE}
glm_database_cont<- genotype_format_filtered %>%
  mutate(sample =  as.numeric(sample)) %>%
  inner_join(phenotypic_data_cont, by = "sample") 

write_rds(glm_database_cont,"intermediate_data/glm_database_cont.rds", "xz")

```

# Validated mutations
We are going to focus on "Key" mutations,i.e. validated mutations reported by the WHO in its last report(Report on antimalarial drug efficacy, resistance and response: 10 years of surveillance (2010-2019)) <https://www.who.int/publications/i/item/9789240012813>
```{r, echo=FALSE}
#We write the list of key mutations
key_mutations<- c(#Reported by WHO k13
                  "F446I", "N458Y", "M476I", "Y493H", "R539T", "I543T", "P553L",
                  "R561H", "P574L", "C580Y", 
                  #candidates K13
                  "P441L", "G449A", "C469F", "C469Y", "A481V", "R515K", "P527H",
                  "N537I", "N537D", "G538V",  "V568G", "R622I", "A675V",
                  #4 aminoquinolines
                  #Pfcrt
                  "K76T", "C72S", "M74I", "N75E", 
                  "A220S","Q271E", "N326S", "I356T", "R371I",
                  #Pfmdr1 in combination with Pfcrt 
                  "N86Y", "Y184F", "S1034C", 'N1042D', "D1246Y", 
                  #Pfcrt
                  "T93S", "H97Y","F145I", "I218F", "C350R", "M343L", "G353V",
                  #Antifolates
                  #PfDHFR
                  "N51I", "C59R", "S108N", "I164L", "A16V",
                  #Pfdhps
                  "S436A", "S436F", "A437G", "K540E", "A581G", "A613T", "A613S",
                  #Pfcytb
                  "Y268N", "Y268S", "Y268C",
                  
                  #published by Verity et al
                  "S160N", "K199T", "I723V", "G639D", "N569K", "E431K", "L402V")

key_mutations_all_plates<- data_all_plates %>%
  filter(aa_change %in% key_mutations, 
gene =="dhps" | gene =="dhfr-ts"| gene=="crt"|gene=="mdr1" )
                     

```

## Frequency of key mutations
```{r, echo=FALSE}
frequency_key_mutations <- mutation_prevalence(key_mutations_all_plates, threshold = 0) 

Key_mutations_plot<- frequency_key_mutations %>%
  mutate(mutation_name2 = mutation_name,
         frequency = num((prevalence*100),digits = 1),
         mutation_name = str_replace(mutation_name,"ts-","")) %>%
 separate_wider_delim(mutation_name2, "-", names = c("gene", "mutation"), too_many = "merge") %>%
  mutate(mutation = str_replace(mutation,"ts-","")) %>%
  mutate(aa_coden_position = as.numeric(str_extract(mutation, "[1-9][0-9]{0,3}"))) %>%
  group_by(gene) %>%
  arrange(aa_coden_position,.by_group = TRUE) %>% 
  ungroup() %>%
  mutate(n_order = row_number()) %>%
  ggplot(aes(reorder(mutation_name, n_order), frequency, fill =  gene)) + 
  geom_bar(stat = "identity") + 
  theme_classic2(base_size = 16) + 
  ylim(0,100)+
  ylab("Frequency (%)") + xlab("Amino acid change") +  ggtitle("") +
  scale_fill_brewer(name = "Gene", palette = "Dark2") +
  theme(plot.title = element_text(hjust = 0.5),
       axis.text.x = element_text(angle=90, vjust=0.5, size = 12), 
        axis.text.y = element_text(angle=0, vjust=0.5, size = 12), 
        legend.position ="bottom") 
           
print(Key_mutations_plot)
ggsave("figures/mut_key_frequency.pdf", units="cm", width=19, height=14, dpi=300)

```

```{r}
key_mutations_prev<-Key_mutations_plot$data %>%
  mutate(frequency = as.numeric(frequency)) 

write.csv(key_mutations_prev, "tables/Key_mutations_prev.csv", row.names = FALSE)

```


# K13 mutations
```{r, message=FALSE, echo=FALSE}
#We use the glm_database_cont database to be able to include all mutations in K13 gene, not just the validated ones
k13 <- glm_database_cont%>%
     select(sample,starts_with("k13"))%>%
  pivot_longer(!c(sample), names_to = "mutation", values_to = "genotype") %>%
  filter(!str_detect(mutation, "del"),
         !str_detect(mutation, "ins"),
         !str_detect(mutation, "fs"),
         !str_detect(mutation, "dup")) %>%
  filter(!is.na(genotype)) %>% #we are including both monoclonal (homozygous , i.e. 2) and polyclonal (heterozygous, i.e. 1) and wild type monoclonal (homozygous, i.e. o)
  mutate(genotype = ifelse(genotype>=1,1,0),
         mutation = str_replace_all(mutation,"k13-","")) %>%
  group_by(mutation) %>%
  #create new variables to count the total number of genotyped samples, the ones that are mutant and the frequency
  mutate(n_total = length(mutation),
         n_mutant = sum(genotype==1),
    frequency = num((n_mutant*100/n_total),digits = 1)) %>%
  select(!c(sample,genotype)) %>%
  unique() %>%
  mutate(aa_coden_position = as.numeric(str_extract(mutation, "[1-9][0-9]{0,3}")),
         Within_propeller_domain = ifelse(aa_coden_position<350,"N-ter","Propeller")) %>%
  arrange(aa_coden_position) %>% 
  ggplot(aes(reorder(mutation, aa_coden_position), frequency, fill = Within_propeller_domain)) + 
  geom_bar(stat = "identity") + 
  theme_classic2(base_size = 12.3) + 
  ylab("Frequency (%)") + xlab("") + 
  scale_fill_brewer(name = "WHO category", palette = "Dark2") +
  theme(plot.title = element_text(hjust = 0.5),
       axis.text.x = element_text(angle=90, vjust=0.5, size = 9), 
        axis.text.y = element_text(angle=0, vjust=0.5, size = 9), 
        legend.position ="none") 
  
print(k13)
 # Validated #F446I N458Y M476I Y493H R539T I543T P553L R561H P574L C580Y

 # Candidate "P441L", "G449A", "C469F", "C469Y", "A481V", "R515K", "P527H",
 #                  "N537I", "N537D", "G538V",  "V568G", "R622I", "A675V",
ggsave("figures/k13_mutations.pdf", units="cm", width=15, height=9, dpi=300)
```
```{r}
K13_frequency<- k13$data %>%
  mutate(frequency = as.numeric(frequency)) %>%
  filter(Within_propeller_domain=="Propeller")

write.csv(K13_frequency, "tables/K13_frequency.csv", row.names = FALSE)

```

*Background mutations*
Pf apicoplast ribosomal protein s10 (Pfarps10-V127M) not covered
Pf ferredoxin (Pffd-D193Y) ok
Pf multidrug resistance 2 transporter (Pfmdr2-T484I) ok
putative phosphoinositide-binding protein (Pfpib7-C1484F) ok
Pf protein phosphatase (Pfpp-V1157L) ok
Pf chloroquine resistance transporter (Pfcrt-N326S) and Pfcrt-I356T), OK


```{r, echo=FALSE, message=FALSE}
background_mutations<-
glm_database_cont%>%
  select(sample,
         `fd-Asp193Tyr`,
         `mdr2-Thr484Ile`,
         `pib7-Cys1484Phe`,
         `crt-Asn326Ser`,
         `crt-Ile356Thr`,
         `pph-Val1157Leu`) %>%
  pivot_longer(!c(sample), names_to = "mutation", values_to = "genotype") %>%
select(mutation,genotype)%>%
   filter(!is.na(genotype)) %>% 
  #we are including both monoclonal (homozygous , i.e. 2) and polyclonal (heterozygous, i.e. 1) and wild type monoclonal (homozygous, i.e. o)
  mutate(genotype = ifelse(genotype>=1,1,0))%>%
  group_by(mutation) %>%
  #create new variables to count the total number of genotyped samples, the ones that are mutant and the frequency
  mutate(n_total = length(mutation), 
         n_mutant = sum(genotype==1),
    frequency = num((n_mutant*100/n_total),digits = 1)) %>%
  select(!c(genotype)) %>%
  unique() 

#In order to preserve numeric columns as classes when exported, the easiest way is to export it to excel file
write_xlsx(background_mutations, "tables/background_mutations.xlsx")


```

*Other mutations* 
Pfap2mu S160N ok

Pfatp6:
L263E, not covered 
E431K, ok
A623E, ok
S769N, ok

Pfubp1:
D1525E, ok
E1528D, ok

```{r, echo=FALSE, message=FALSE}
other_mutations<-
glm_database_cont %>%
  select(sample,
         `Pfap2mu-Ser160Asn`,
         `atp6-Glu431Lys`,
         `atp6-Ala623Glu`,
         `atp6-Ser769Asn`,
         `Pfubp1-Asp1525Glu`,
         `Pfubp1-Glu1528Asp`) %>%
pivot_longer(!c(sample), names_to = "mutation", values_to = "genotype") %>%
select(mutation,genotype)%>%
  filter(!is.na(genotype)) %>%
  mutate(genotype = ifelse(genotype>=1,1,0)) %>%
  group_by(mutation) %>%
  #create new variables to count the total number of genotyped samples, the ones that are mutant and the frequency
  mutate(n_total = length(mutation), 
         n_mutant = sum(genotype==1),
    frequency = num((n_mutant*100/n_total),digits = 1)) %>%
  select(!c(genotype)) %>%
  unique() 

 
#In order to preserve numeric columns as classes when exported, the easiest way is to export it to excel file
write_xlsx(other_mutations, "tables/other_mutations.xlsx")

```

## Frequency table per year 
```{r}
IC50_genotype_subset<-
                    IC50_genotype %>%
                    #We select key genes
               select(starts_with("dhps"),
                      starts_with("dhfr"),
                      starts_with("crt"),
                      starts_with("mdr1"),
                      starts_with("k13"),
                      year,
                      country, sample) %>%
  pivot_longer(!c(sample,country, year), names_to = "mutation", values_to = "genotype") %>%
  filter(!genotype== "-1") %>% #we are including both monoclonal (homozygous , i.e. 2) and polyclonal (heterozygous, i.e. 1) and wild type monoclonal (homozygous, i.e. o)
  mutate(genotype = ifelse(genotype>=1,1,0)) %>%
           separate_wider_delim(mutation, "-", names = c("gene", "mutation"), too_many = "merge") %>%
  mutate(mutation = str_replace(mutation,"ts-",""),
         gene = str_replace(gene,"dhfr","dhfr-ts"))
```

```{r}

IC50_genotype_subset %>%
  filter(mutation %in% key_mutations_prev$mutation | 
           mutation %in% K13_frequency$mutation) %>%
  select(year,gene,mutation,genotype) %>%
  mutate(year = as.character(year),
  year = case_when(year=="2020" ~ "2020-21", 
                          year=="2021"  ~ "2020-21", 
                        year=="2022" ~ "2022-23", 
                          year=="2023" ~ "2022-23", 
                          TRUE~ year)) %>% 
  group_by(gene,mutation,year) %>%
  #create new variables to count the total number of genotyped samples, the ones that are mutant and the frequency
  mutate(n_total = length(mutation),
         n_mutant = sum(genotype==1),
    frequency = num((n_mutant*100/n_total),digits = 1)) %>%
  select(!c(genotype, n_total, n_mutant)) %>%
  unique() %>%
  mutate(aa_coden_position = as.numeric(str_extract(mutation, "[1-9][0-9]{0,3}"))) %>%
  group_by(gene) %>%
  arrange(aa_coden_position,.by_group = TRUE) %>% 
  ungroup() %>%
  mutate(n_order = row_number()) %>%
  select(!c(aa_coden_position, n_order)) %>%
  pivot_wider(names_from = year, values_from = frequency) %>%
  DT::datatable(
    extensions = 'Buttons',
    options = list(dom = 'Bfrtip',
                   buttons = c('excel', 'csv'))
  )

```




```{r, echo=FALSE, message=FALSE}

top_countries <- IC50_genotype %>%
  mutate(country= case_when(country== "Congo_RDC_Zaire"~ "Democratic Republic of the Congo",
                                  country == "Congo"~ "Republic of the Congo",
                              country == "Guinee Equatoriale"  ~ "Equatorial Guinea",
                             country == "Comores" ~ "Comoros",
                             country == "Central Africa" ~ "Central African Republic", 
                                  TRUE ~ country)) %>%
  rename(admin = country) %>%
  group_by(admin) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  mutate(Frequency = n/sum(n)) %>%
  arrange(desc(Frequency)) %>%
  filter(Frequency>=0.026) %>%
  mutate(admin2 = admin) %>%
  mutate(admin2 = case_when(admin2=="Benin" |admin2=="Gabon" |admin2=="Togo"|admin2=="Burkina Faso" |admin2=="Nigeria"|admin2=="Democratic Republic of the Congo" | admin2=="Niger" ~ "agreggate", TRUE ~ admin2))

```

# "Raw data" i.e. categorized into wild-type, mixed genotype and pure mutant
```{r, echo=FALSE, message=FALSE}

IC50_genotype_subset_raw<- IC50_genotype %>%
  #We select key genes
               select(starts_with("dhps"),
                      starts_with("dhfr"),
                      starts_with("crt"),
                      starts_with("mdr1"),
                      starts_with("k13"),
                      year,
                      country, sample) %>%
  pivot_longer(!c(sample,country, year), names_to = "mutation", values_to = "genotype") %>%
  filter(!genotype== "-1") %>% #we are including both monoclonal (homozygous , i.e. 2) and polyclonal (heterozygous, i.e. 1) and wild type monoclonal (homozygous, i.e. o)
           separate_wider_delim(mutation, "-", names = c("gene", "mutation"), too_many = "merge") %>%
  mutate(mutation = str_replace(mutation,"ts-",""),
         gene = str_replace(gene,"dhfr","dhfr-ts"))


key_genotype_subset_raw<-
IC50_genotype_subset_raw %>%
  filter(mutation %in% key_mutations_prev$mutation | 
           mutation %in% K13_frequency$mutation) %>%
  mutate(year = as.character(year),
  year = case_when(year=="2020" ~ "2020-21", 
                          year=="2021"  ~ "2020-21", 
                        year=="2022" ~ "2022-23", 
                          year=="2023" ~ "2022-23", 
                          TRUE~ year)) %>% 
  mutate(genotype = case_when(genotype==1 ~ "mixed",
                              genotype==0 ~ "wild-type",
                              genotype==2 ~ "mutant")) %>%
  group_by(gene,mutation,year) %>%
   mutate(aa_coden_position = as.numeric(str_extract(mutation, "[1-9][0-9]{0,3}"))) %>%
  group_by(gene) %>%
  arrange(aa_coden_position,.by_group = TRUE) %>% 
  ungroup() %>%
  mutate(n_order = row_number()) %>%
  select(!c(aa_coden_position, n_order)) %>%
  mutate(gene_mut = paste(gene,mutation, sep = "-"),
         genotype = as.character(genotype)) %>%
   pivot_wider(id_cols= c(year, country, sample), names_from = gene_mut, values_from = genotype) 

```

# "Raw data" table per year, i.e. categorized into wild-type, mixed genotype and pure mutant
```{r, echo=FALSE, message=FALSE}
key_genotype_subset_raw_no_country<-
key_genotype_subset_raw %>%
  select(!c(country, sample))

key_genotype_subset_raw_table<- table1(~.| year
     , data = key_genotype_subset_raw_no_country, 
     render.missing = NULL,
     render.categorical="FREQ (PCTnoNA%)"
     )  
knitr::kable(key_genotype_subset_raw_table)

```

```{r,echo=FALSE, message=FALSE}
write.csv(key_genotype_subset_raw_table, "tables/key_genotype_subset_raw_table.csv", col.names = F)
```

# "Raw data" table per top country, i.e. categorized into wild-type, mixed genotype and pure mutant
```{r, echo=FALSE}
key_genotype_subset_raw_country<-
key_genotype_subset_raw %>%
  select(!c(year, sample)) %>%
 filter(country %in% top_countries$admin)

key_genotype_subset_raw_country_table<- table1(~.| country
     , data = key_genotype_subset_raw_country,  
     render.missing = NULL,
     render.categorical="FREQ (PCTnoNA%)" ) 
 
knitr::kable(key_genotype_subset_raw_country_table)
```

```{r,echo=FALSE, message=FALSE}
write.csv(key_genotype_subset_raw_country_table, "tables/key_genotype_subset_raw_country_table.csv", col.names = F)
```

# Temporal change of key resistance mutations by year
```{r, echo=FALSE, message=FALSE}
mutations_trend_time<- IC50_genotype_subset %>%
      filter(mutation == "Lys76Thr" |mutation == "Ala437Gly" | mutation == "Cys59Arg" |
                             mutation == "Asn86Tyr") %>%
  mutate(mutation = case_when(mutation =="Lys76Thr" ~ "Pfcrt (K76T)",
                              mutation =="Ala437Gly" ~ "Pfdhps (A437G)",
                              mutation =="Cys59Arg" ~ "Pfdhfr-ts (C59R)",
                              mutation == "Asn86Tyr" ~ "Pfmdr1 (N86Y)")) %>%
  mutate(year = as.character(year),
  year = case_when(year=="2020" ~ "2020-21", 
                          year=="2021"  ~ "2020-21", 
                        year=="2022" ~ "2022-23", 
                          year=="2023" ~ "2022-23", 
                          TRUE~ year)) %>% 
  group_by(mutation, year) %>%
  #create new variables to count the total number of genotyped samples, the ones that are mutant and the frequency
  mutate(n_total = length(mutation),
         n_mutant = sum(genotype==1),
    frequency = num((n_mutant/n_total),digits = 2),
         m = (qbinom(0.5, size=n_total, prob=frequency))/n_total,
         lim_inf = (qbinom( 0.025, size=n_total, prob=frequency))/n_total,
         lim_sup = (qbinom( 0.975, size=n_total, prob=frequency))/n_total) %>%
  select(!c(genotype, n_total, n_mutant, sample, country)) %>%
  unique() %>%
  ggplot(aes(year, frequency, group = mutation, colour = mutation)) + 
  geom_point(size = 2) +  geom_line() + 
  geom_pointrange(aes(x =year, ymin=lim_inf, ymax=lim_sup),alpha = 0.5 ) +
  scale_colour_brewer(palette = "Set1") + scale_y_continuous(breaks = c(0, 0.25, 0.5,0.75, 1), labels = c(0, 25, 50, 75, 100)) + 
  ylab("Frequency (%)") + xlab ("Year") + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title= element_text(size = 18),
        axis.text.x = element_text(angle=90, vjust=0.5, size = 18),
        axis.text.y = element_text(angle=0, vjust=0.5, size = 18))


print(mutations_trend_time)
ggsave("figures/mutations_trend_time.pdf", units="cm", width=19, height=14, dpi=300)

```
# Frequency of key resistance mutations by year and by country
```{r}
mutations_trend_time_top_ten_countries<- IC50_genotype_subset %>%
      filter(mutation == "Lys76Thr" | 
              mutation == "Ala437Gly" | mutation == "Cys59Arg" |
              mutation == "Asn86Tyr",
             country %in% top_countries$admin) %>%
    mutate(mutation = case_when(mutation =="Lys76Thr" ~ "Pfcrt (K76T)",
                              mutation =="Ala437Gly" ~ "Pfdhps (A437G)",
                              mutation =="Cys59Arg" ~ "Pfdhfr-ts (C59R)",
                              mutation == "Asn86Tyr" ~ "Pfmdr1 (N86Y)")) %>%
  mutate(year = as.character(year),
  year = case_when(year=="2020" ~ "2020-21", 
                          year=="2021"  ~ "2020-21", 
                        year=="2022" ~ "2022-23", 
                          year=="2023" ~ "2022-23", 
                          TRUE~ year)) %>%
 mutate(country = factor(country, 
                           levels=c("Ivory Coast","Cameroon","Mali", "Guinea","Republic of the Congo", "Senegal", "Central African Republic", "Chad", "Gabon", "Benin"),
                           labels = c("Côte d'Ivoire","Cameroon","Mali", "Guinea","Republic of the Congo", "Senegal", "Central African Republic", "Chad", "Gabon", "Benin"))) %>% 
  group_by(mutation, year, country) %>%
  #create new variables to count the total number of genotyped samples, the ones that are mutant and the frequency
  mutate(n_total = length(mutation),
         n_mutant = sum(genotype==1),
    frequency = num((n_mutant/n_total),digits = 2),
         m = (qbinom(0.5, size=n_total, prob=frequency))/n_total,
         lim_inf = (qbinom( 0.025, size=n_total, prob=frequency))/n_total,
         lim_sup = (qbinom( 0.975, size=n_total, prob=frequency))/n_total) %>%
  select(!c(genotype, n_total, n_mutant, sample)) %>%
  unique() %>%
  ggplot(aes(year, frequency, group = mutation, colour = mutation)) + 
  geom_point(size = 2) +  geom_line() + 
  geom_pointrange(aes(x =year, ymin=lim_inf, ymax=lim_sup),alpha = 0.5 ) +
  scale_colour_brewer(palette = "Set1") + scale_y_continuous(breaks = c(0, 0.25, 0.5,0.75, 1), labels = c(0, 25, 50, 75, 100)) + 
  ylab("Frequency (%)") + xlab ("Year") + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))+
  facet_wrap(~country) 
             
print(mutations_trend_time_top_ten_countries)
ggsave("figures/mutations_trend_time_top_ten_countries.pdf", units="cm", width=23, height=14.5, dpi=300)

```

```{r echo=F, eval=F}
save.image("./intermediate_data/validated_mutations.RData")
```

